Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-05-31 measurements.tex
expanded results multicore

The authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the Raspberry Pi 2 Model B, which
features four ARM Cortex-A7 cores in a Broadcom BCM2836 \ac{SOC} with a 900 MHz
clock. Each core has a 32 KiB L1 data cache and 32 KiB L1 instruction cache. The
cores share a 512 KiB L2 cache. All the measured results, including the baseline
results, were obtained with NEON-enabled code adopted from the FIFI/KODO library
\cite{kodo2011pedersen}. The NEON extension provides an 128-bit \ac{SIMD}
instruction set to the Raspberry Pi 2. Figures~\ref{enc_dec1024},
\ref{enc_dec128} and \ref{enc_dec16} show the encoding and decoding throughput
in MiB per second for different generation sizes ($g$ = 1024, 128, and 16
respectively). The throughput is the rate of generating $g$ coded packets,
divided by the time the encoder took to perform the task. The size of each coded
packet was fixed to $1536$~bytes since that is the typical size of an Ethernet
frame. The blocked operations were performed dividing the matrices in sub-blocks
of 16,32,64,...,1024 elements, and the figures only show the best results (usually obtained with a block size of 16 or 32 since with bigger block sizes, the operands do not fit in L1 cache).

@@

\begin{figure}[ht!]
\centering
\includegraphics[width=\columnwidth]{images/2015-04-18_encoding_decoding_1024.eps}
\caption{Encoding and Decoding performance for g = 1024}
\label{enc_dec1024}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[width=\columnwidth]{images/2015-04-18_encoding_decoding_128.eps}
\caption{Encoding and Decoding performance for g = 128}
\label{enc_dec128}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[width=\columnwidth]{images/2015-04-18_encoding_decoding_16.eps}
\caption{Encoding and Decoding performance for g = 16}
\label{enc_dec16}
\end{figure}

@@

>>>>

The authors implemented the algorithm described in
Section~\ref{sub:implementation-multicore} on the Raspberry Pi 2 Model B, which
features four ARM Cortex-A7 cores in a Broadcom BCM2836 \ac{SOC} with a 900 MHz
clock. Each core has a 32 KiB L1 data cache and 32 KiB L1 instruction cache. The
cores share a 512 KiB L2 cache. All the measured results, including the baseline
results, were obtained with NEON-enabled code adopted from the FIFI/KODO library
\cite{kodo2011pedersen}. The NEON extension provides an 128-bit \ac{SIMD}
instruction set to the Raspberry Pi 2. Figures~\ref{enc_dec1024},
\ref{enc_dec128} and \ref{enc_dec16} show the encoding and decoding throughput
in MiB per second for different generation sizes ($g$ = 1024, 128, and 16
respectively). The throughput is the rate of generating $g$ coded packets,
divided by the time the encoder took to perform the task. The size of each coded
packet was fixed to $1536$~bytes since that is the typical size of an Ethernet
frame. The blocked operations were performed dividing the matrices in sub-blocks
of 16,32,64,...,1024 symbols, and the figures only show the best results
(usually block sizes of 16 or 32 symbols since with bigger block sizes, the
operands do not fit in L1 cache).

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{images/2015-04-18_encoding_decoding_1024.eps}
\caption{Encoding and Decoding performance for g = 1024}
\label{enc_dec1024}
\end{figure}

For $g=$ 1024, the blocked baseline measurements outperforms the non blocked
variant. This means that making the matrix multiplication algorithm cache
efficient brings an increase in throughput by a factor of 2.83. When using the
algorithm described in Section~\ref{sub:implementation-multicore}, encoding with
four cores is on average 3.7 times faster than with one core. Similarly,
decoding with four codes is 3.05 times faster, on average, than decoding with a
single core. Figure~\ref{enc_dec1024} shows that the implemented algorithm, by
exploiting cache efficiency and only three extra cores provides a gain of 10
folds compared with traditional non-blocked algorithms. With $g$ = 1024, the
matrix inversion becomes more expensive than at smaller generations sizes.
Therefore, the decoding throughput is 46\% of the encoding throughput.

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{images/2015-04-18_encoding_decoding_128.eps}
\caption{Encoding and Decoding performance for g = 128}
\label{enc_dec128}
\end{figure}

For $g$ = 128, the differences between the baselines operations show that a
blocked algorithm is 15\% faster than the non-blocked variant. Encoding with
four cores is 2.82 times faster than with a single core. Due to the smaller
matrix sizes, the gain by using blocked operations in the baselines is not that
big compared with $g$ = 1024. For the same reason, the matrix inversion is less
expensive. As a consequence, the decoding throughput is 71\% of the encoding
throughput.

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{images/2015-04-18_encoding_decoding_16.eps}
\caption{Encoding and Decoding performance for g = 16}
\label{enc_dec16}
\end{figure}

When $g$ = 16, the gains of blocked operations are negligible compared with the
non-blocked ones. The reason behind this behavior is that all the data fits in
the L1 cache. For the scheduled version, since the problem to solve is so small,
the gain of using four cores is a factor of 1.68 compared with a single core,
and 1.01 compared with two cores. Therefore, there are no practical benefits in
using four cores instead of two.

The differences in throughput, for all generation sizes, between the blocked
baseline and the scheduled measurements are due the time spent resolving the
dependencies and the scheduling overhead. These effects are negligible for big
generation sizes, while considerable for small matrices.

% \begin{figure}
% \centering
% \begin{subfigure}[b]{0.3\textwidth}
% \includegraphics[width=\textwidth]{images/2015-04-18_encoding_decoding_1024.eps}
% \caption{Encoding and Decoding performance for g = 1024}
% \label{enc_dec1024}
% \end{subfigure}
%
% \begin{subfigure}[b]{0.3\textwidth}
% \includegraphics[width=\textwidth]{images/2015-04-18_encoding_decoding_128.eps}
% \caption{Encoding and Decoding performance for g = 128}
% \label{enc_dec128}
% \end{subfigure}
%
% \begin{subfigure}[b]{0.3\textwidth}
% \includegraphics[width=\textwidth]{images/2015-04-18_encoding_decoding_16.eps}
% \caption{Encoding and Decoding performance for g = 16}
% \label{enc_dec16}
% \end{subfigure}
% \end{figure}

@@

To compare how much slower is the matrix multiplication with the matrix
inversion for different generation sizes, we ran a set of tests. We used a
single core to perform the operations. We changed the generation sizes,
performed matrices multiplications and matrix inversions, and measured the spent
time doing so. We then calculated the ratio $ratio = \frac{T_{mult}}{T_{inv}}$.
Table~\ref{runtimes} summarizes the results. The bigger the matrix size, the
smaller is the calculated ratio. This means that when the problems are bigger,
the decoding throughput decreases compared with the encoding throughput.
